---
title: "Project Instructions"
author: "Courtney Cooper"
date: "4/25/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Order of Files-

1. bt maize cleaning.rmd (reading in raw data files and cleaning to produce rmd file)
2. bt maize analysis.rmd (analysis/regressions run from rmd file)
3. whiteanalysis.r (analysis done of just white maize)

packages needed:

```{r packages}


install.packages("pacman")
library(pacman)

p_load(ggplot2, tidyr, dplyr)

install.packages("gt")
install.packages("RColorBrewer")
install.packages("kableExtra")
install.packages("modelsummary")
install.packages("gtsummary")
install.packages("fixest")
library(fixest)

library(gt)
library(sandwich,
        lmtest,
        gtsummary,
        kableExtra)
library(modelsummary)
library(RColorBrewer)
library(modelsummary)
library(stringr)

install.packages("devtools")
devtools::install_github("jaredhuling/jcolors")
library(jcolors)


install.packages("maptools", repos="http://R-Forge.R-project.org")
library(maptools)




```

reg1 <- feglm(yield ~ GM + color | provence + factor(year)+ irrigated, dat)
summary(reg1)

dat$yearsq <- dat$year*dat$year

reg2 <- feglm(yield ~ GM + color + i(year*GM) + i(yearsq*GM) | provence + factor(year)+ irrigated, dat)
summary(reg2)


dat$y_effect <- reg2$coefficients["GM"] + reg2$coefficients["GM:year"] * dat$year
dat$ysq_effect <- reg2$coefficients["GM"] + reg2$coefficients["GM:year"] * dat$year + reg2$coefficients["GM:yearsq"] * dat$yearsq

max(dat$ysq_effect)


# Provence by year by GM effects in one model

## LEFT OFF HERE ** above
reg3 <- feglm(yield ~ GM + color+ i(year*provence*GM) + i(provence*yearsq*GM) | provence + factor(year)+ + irrigated, dat)
summary(reg3)

models <- list(reg1,reg2,reg3)


modelregress <- msummary(models, vcov = vcovHC, stars = TRUE)
print(modelregress)
reduced <- c('(Intercept)' = 'Constant',
             'provenceFS' = "Free State",
             'provenceGP' = "GP",
             'provenceKZN' = "KZN",
             'provenceMP' = "MP",
             'provenceNC' = "NC",
             'provenceNW' = "NW",
             'provenceWS' = "WS",
             'GM' = "GM",
             'coloryellow' = "Color (yellow)",
             'irrigated' = "Irrigated",
             'GM x year' = "GM x year",
             'GM x yearsq' = "GM x yearsq"
             )

msummary(modelregress, coef_map = reduced)

